Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-02-11 reports/paper-rda.tex
Updated the paper to include a lot of the theory based on Ye and Wang (2006). Also, reorganized a lot of the paper to be more coherent.

\title{Another Look at Regularized Discriminant Analysis for High-dimensional Classification}

@@

\newtheorem*{thm}{Theorem}
\newtheorem*{cor}{Corollary}

@@

\begin{abstract}
\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which employs a ridge-like shrinkage factor to ensure positive definiteness. We demonstrate that the \emph{RDA} classifier, in its original form, is inadequate for the small-sample, high-dimensional setting. We redefine the \emph{RDA} classifier to increase its computational runtime and its interpretation. We show that this new definition includes several well-known regularized classifiers as special cases. We also consider several properties of the newly defined \emph{RDA} classifier. Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.  Because the eigenvalues are often highly skewed, this average is often excessively large, so that the near-zero eigenvalues are substantially overcompensated. A comparable eigenvalue adjustment method from \cite{Thomaz:2006ef} often yields a shrinkage factor that is approximately equal to the average of the eigenvalues. By writing each shrinkage method as a matrix function, we generalize the shrinkage method of \cite{Thomaz:2006ef} to allow a powerful estimation technique that does not over inflate the near-zero eigenvalues. By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier. Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and three small-sample, high-dimensional microarray data sets. Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

@@

In our first classifier we generalize the eigenvalue replacement employed in the \emph{NLDA} classifier and apply a soft threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator. We refer to our first classifier as the \emph{RDA} with soft eigenvalue-thresholding (\emph{Soft-ETRDA}) classifier.  As we show, the eigenvalue replacement employed in the \emph{NLDA} classifier is often approximately equal to the shrinkage applied in the \emph{RDA} classifier for small-sample, high-dimensional data. Furthermore, a special case of the \emph{Soft-ETRDA} classifier is the \emph{NLDA} classifier as well as the \emph{RDA} classifier.  We remark that our soft eigenvalue threshold is similar to and partially motivated by the shrunken centroids classifier from \cite*{Tibshirani:2002ht}. 	In our second proposed classifier, we apply a hard threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator, so that the number of covariance parameters incorporated into the \emph{RDA} classifier is reduced substantially. We refer to this classifier as the \emph{RDA} with hard eigenvalue-thresholding (\emph{Hard-ETRDA}) classifier. We remark that the inverse of \emph{Hard-ETRDA} covariance-matrix estimator includes as a special case, the Moore-Penrose pseudo inverse of the \emph{RDA} covariance-matrix estimator.

@@

We have organized the remainder of the paper as follows. In Section 2 we review the \emph{RDA}, the \emph{MDEB}, and \emph{NLDA} classifiers and define each in terms of matrix functions. In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers. In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 6.

@@

We\footnote{Define the majority of notation here, including $z^+$, $z_+$, and $I[z]$.} wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
\begin{align}
	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
\end{align}
where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}

@@

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.

@@

Substituting \eqref{eq:sig-rda} into \eqref{eq:qda}, we have the \emph{RDA} classifier, where we assign an unlabeled observation $\bm x$ to class $\omega_k$:
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}

\subsection{An Alternative Approach to Regularized Discriminant Analysis}

Here, we present a more straightforward, alternative parameterization of the \emph{RDA} classifier. We first define explicitly
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. Before examining the shrinkage applied to \eqref{eq:sig-lambda-alternative}, we examine some of its properties and write \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'}.\label{eq:sig-lambda-alternative2}
\end{align}
From \eqref{eq:sig-lambda-alternative2}, we see the weight $\lambda$ places on the covariance matrices from the remaining $K - 1$ classes. The interesting part here is that $\lambda$ controls the contribution of each of the $N$ observations to estimating $\Sigma_k$. Effectively, this states that for $0 < \lambda \le 1$, we use $N$ observations to estimate $\bm \Sigma_k$ rather than $n_k$ as done in the \emph{QDA} classifier. As we will discuss in-depth later, the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from $n_k - 1$ to $N - 1$ for $0 < \lambda \le 1$. Moreover, suppose that we have centered each observation $\bm x_i$, $i = 1, \ldots, N$, by its class sample mean. Then, we write \eqref{eq:sig-lambda-alternative2} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative3}
\intertext{where}
	c_{ik}(\lambda) &= \begin{cases}
		(1 - \lambda)n_k^{-1} + \lambda N^{-1}, & y_i = k,\\
		\lambda N^{-1}, & y_i \ne k.\\
	\end{cases}\nonumber
\end{align}
Therefore, after centering each observation by its corresponding class sample mean, we have that $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda-alternative3} incorporates each centered observation into the estimation of $\bm \Sigma_k$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative3} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$. Let $e_{jk}(\lambda)$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}(\lambda)$ be the corresponding eigenvector such that $e_{1k}(\lambda) \ge e_{2k}(\lambda) \ge \ldots \ge e_{pk}(\lambda) \ge 0$. Equivalently, we write \eqref{eq:qda} as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}(\lambda)' (\bm x - \bm \xbar_k)]^2}{e_{jk}(\lambda)} + \sum_{j=1}^p \ln \{e_{jk}(\lambda)\}. \label{eq:rda-eigen}
\end{align}
The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:rda-eigen}. In fact, the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}. This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k(\lambda)) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are 0, which implies that \eqref{eq:rda-eigen} is incalculable. Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to stabilize the eigenvalues of $\widehat{\bm \Sigma}_k$ used in \eqref{eq:qda}. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform \citep{Ch15SSLBook} or a transfer function \citep{Ye:2009gd, Chapelle2002SSLChapter15Reference} . Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

>>>>

\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}

@@

\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{thm}{Theorem}

@@

\begin{abstract}
\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness. We show that this shrinkage is unnecessary because shrinking the zero eigenvalues corresponding to the null space of the pooled sample covariance matrix has no effect on the \emph{RDA} classifier. Accordingly, we propose a novel generalization of the \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes. Additionally, our formulation of the \emph{RDA} classifier allows us to estimate efficiently optimal tuning parameters from a large set of candidate values. Furthermore, our generalization simplifies the construction of the covariance-matrix estimator to improve its interpretation and allows for a more flexible classification model by viewing shrinkage applied to the nonzero eigenvalue estimates in terms of matrix functions. Fortuitously, this matrix-function formulation includes as special cases the original \emph{RDA} classifier and several other well-known regularized classifiers. Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers. Finally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

@@

@@

Quote from \cite{Ye:2009gd}: ``There is a lack of a systematic study to explore the commonalities and differences of these algorithms, as well as their intrinsic relation- ship. This has been a challenging task, since different algorithms apply completely different schemes when dealing with the singularity problem.'' Cite this as an issue with LDA and that regularization has been less explored with QDA. Our formulation provides the necessary tools to study this is in detail.

\cite{Ye:2006jm} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier when $\lambda = 1$. Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits. First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$. Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier. For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}

As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.\footnote{TODO: Contrast the computational efficiency we gain with the need to stabilize the smallest eigenvalues (those near $q$).}

We prove several results using arguments similar to \cite{Ye:2006tq}.

We have organized the remainder of the paper as follows. In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}. In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them. In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 5.

@@

In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is applied instead for each class, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
In this case \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier, and the log-determinant is omitted because it is constant across all classes.

The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, and this bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Therefore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$ be monotonic. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform \citep{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}. Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, no eigenvalue adjustment is performed, and \eqref{eq:qda} remains incalculable. Similarly, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, the \emph{LDA} classifier is employed without eigenvalue shrinkage. Additionally, setting $\phi(e_j) = e_j^{+}$ with $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise, we ignore the zero eigenvalues, so that $\phi(\widehat{\bm \Sigma}_k) = \widehat{\bm \Sigma}_k^{+}$, the Moore-Penrose pseudoinverse of $\widehat{\bm \Sigma}_k$.

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}) = e_{jk} + \gamma_k, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression. Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite. For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.

@@

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.

@@

\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to obtain the \emph{RDA} classifier
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
Notice that we can write the shrinkage term in \eqref{eq:sig-rda} as the matrix function
\begin{align*}
	\phi(e_{jk}(\lambda)) = (1 - \gamma) e_{jk}(\lambda) + \gamma \bar{e}_k,
\end{align*}
for class $\omega_k$, where $e_{jk}(\lambda)$ is the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}(\lambda)$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.

\subsection{A Generalization of Regularized Discriminant Analysis}

Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means. Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation. This formulation allows us to adjust the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in a more general manner so that we are not limited to ridge-like estimators.
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations. Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma}_k)$ for $0 < \lambda \le 1$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0. Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$. To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$. Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
\begin{align}
	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
\end{align}

\section{Squiggles}
We require the following relationship about the null spaces of $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ in order to attain improved computational efficiency of our generalized \emph{RDA} classifier.

\begin{lemma}\label{lemma:null-spaces}
Let $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ be defined as above. Then, $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
\end{lemma}
\begin{proof}
Let $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k(\lambda))$ for some $k = 1, \ldots, K$. Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 - \lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ are each positive semidefinite, we have $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. Now, suppose $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$. We have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ once again because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
\end{proof}

Here, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ to decompose the decision rule of our proposed classifier. In particular, the alternative expression is terms of the first $q$ regularized eigenvalues of $\tilde{\bm \Sigma}_k$ and the adjustments applied to the $p - q$ zero eigenvalues.
